---
title: "Problem11.9"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
horse<-read_csv("https://instruction.bus.wisc.edu/jfrees/jfreesbooks/Regression%20Modeling/BookWebDec2010/CSVData/HKHorse.csv")
```

```{r}
horse %>%
  ggplot(aes(x=WIN, y=FINISH))+
  geom_point()+
  scale_x_continuous(name="WIN %", limits=c(0,1))
```

### c.)

```{r}
horse%>%group_by(FINISH)%>%
  summarise(min=summary(WIN)[1],
            Q1=summary(WIN)[2],
            median=summary(WIN)[3],
            Q3=summary(WIN)[5],
            max=summary(WIN)[6],
            avg=summary(WIN)[4],
            sd=sd(WIN),
            num=length(WIN)
            )
```
Because there are less horses that win the race, their averages are a bit inflated. With the horses that lose, they were already predicted to lose more often than win.

### d.)
```{r}
horseLM <- lm(FINISH~WIN, data=horse)
summary(horseLM)
```

#### i.)
From the summary above, we can see that WIN is a significant predictor to the FINISH variable.

#### ii.)
$$H_O: \text{The linear model does not fits the data well}$$

$$H_A: \text{The linear model does fits the data well}$$

```{r}
testStat <- sum((horse$FINISH-fitted.values(horseLM))^2/fitted.values(horseLM))
testStat
1-pchisq(testStat, length(horse$FINISH)-2)
```
Since the p-value from the chi-square test of 0.9758 is greater than even an $\alpha=0.1$. We fail to reject the null hypothesis. This means there is no statistically significant evidence that the linear model fits the data well.

#### iii.)
It is not possible for the fitted values to lie outside the interval (0,1]. Since the range of fitted values are between [0.008, 0.884].

### e.)
```{r}
horseGLM <- glm(FINISH~WIN, family=binomial(link="logit"), data=horse)
summary(horseGLM)
```
WIN is still a significant predictor for a horse Winning the race.

#### i.)
```{r}
pointEstimates <- cbind(rep(1, times=5), c(0, 0.01, 0.05, 0.10, 1))
fittedLM <- pointEstimates %*% coef(horseLM)
fittedLM
fittedGLM <- exp(pointEstimates %*% coef(horseGLM))/(1+exp(pointEstimates %*% coef(horseGLM)))
fittedGLM
```

#### ii.)
```{r}
as_tibble(cbind(fitted.values(horseLM), fitted.values(horseGLM))) %>%
  ggplot(aes(x=V1, y=V2))+
  geom_point()+
  scale_x_continuous(name="Fitted values from LM", limit=c(0,1))+
  scale_y_continuous(name="Fitted values from GLM", limit=c(0,1))
```

```{r}
fittedLM2 <- predict(horseLM)
plot(fittedLM2, horse$FINISH)
fittedGLM2 <- predict(horseGLM, type="response")
plot(fittedGLM2, horse$FINISH)
```