---
title: "Problem3.6"
output: html_document
---

```{r include=FALSE}
library(tidyverse);
library(rmarkdown);
library(knitr)
```
### Importing and Setting up the Data
```{r}
UNLifeExpectancy <- read.csv(file.path(str_remove(getwd(), "/Class Notes"), "/Book Data/UNLifeExpectancy.csv"))
view(UNLifeExpectancy)
UNLifeExpectancy <- UNLifeExpectancy %>%
  mutate(lnHEALTH = log(PRIVATEHEALTH))
```
### Generating the Regression Model
```{r}
LifeExpReg <- lm(LIFEEXP ~ FERTILITY+PUBLICEDUCATION+lnHEALTH, data = UNLifeExpectancy)
```

### Summarizing the Regression Model

```{r}
summary(LifeExpReg)
```

$$\widehat{LifeExp}=85.6264-5.3993x_{Fertility}-0.1846x_{PublicEducation}-1.0296x_{logHealth}$$

### Problems

#### a.)

Holding all other explanatory variables constant. For every 1% increase of a country's GDP being spent on public education, we can expect the average life expectancy of that country to decrease by ~0.2 years.

#### b.)

Holding all other explanatory variables constant. For every factor of $e$ increase in the country's health expenditures, we can expect the average life expectancy of that country to decrease by ~1 year.

#### c.)
$$
H_O:\beta_{PublicEducation}=0 \newline
H_A:\beta_{PublicEducation}\not=0
$$
If the minimum p-value from the 2-sided t-test is greater than $\alpha$=5%, then we fail to reject $H_O$. \
However, if the minimum p-value from the 2-sided t-test is less than $\alpha$=5%, then we reject $H_O$\
From the summary of the LifeExpReg, we can see that the p-value from the 2-sided t-test for Public Education was 49.3%. Since our p-value of 49.3% is greater than $\alpha$=5%, we fail to reject the null hypothesis. Therefore, there is no statistically significant evidence that $\beta_{PublicEducation}$ is different from 0.

Here is the scatter plot of Public Education Expenditure compared to Life Expectancy:
<center>
```{r echo=FALSE, warning=FALSE}
UNLifeExpectancy %>%
  ggplot(aes(x=PUBLICEDUCATION, y=LIFEEXP)) +
  geom_point() +
  labs(title= "Scatterplot between Life Expectancy and Public Education Expenditure",
       x= "Public Education Expenditure in %",
       y= "Life Expectancy in years") +
  theme_bw()
```
</center>


### 3.3

```{r}
Dat <- matrix(c(-1, 2, 4, 6,
                0, 0, 1, 1,
                0, 1, 5, 8),
              byrow = TRUE, nrow=3, ncol=4,
              dimnames=list(c("xi1", "xi2", "yi"), c(1:4)))
b <- c(0.15, 0.692, 2.88)
H <- matrix(c(0.53846, -0.07692, -0.15385,
              -0.07692, 0.15385, -0.69231,
              -0.15385, -0.69231, 4.11538),
            byrow=TRUE, nrow=3, ncol=3)
s <- 1.373
```

#### a.)
```{r}
y <- Dat[3,]
X <- rbind(rep(1, times=4),Dat[1:2,])
```

#### b.)
```{r}
t(X) %*% b
```

$\hat{y}_3=5.798$

#### c.)
```{r}
H
s * sqrt(4.11538)
```
$se(b_2)=2.785323$