---
title: "Problem4.4"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(tidyverse)
```

### Initializing the Data
```{r}
AutoClaims <- read.csv(file.path(str_remove(getwd(), "/Class Notes"), "/Book Data/AutoClaims.csv"))
AutoClaims <- AutoClaims %>%
  mutate(LNPAID = log(PAID))
```

### Problem 4.4

#### a.)
```{r}
AutoClaimsRega <- lm(LNPAID ~ AGE, data = AutoClaims)

summary(AutoClaimsRega)
```

$$
H_0: \beta_{Age} = 0 \\
H_A: \beta_{Age} \not= 0
$$
If the minimum p-value from the two-sided t-test for $\beta_{Age}$ is less than $\alpha=0.05$, then we have statistically significant evidence to reject the null hypothesis.

Since the minimum p-value from the two-sided t-test for $\beta_{Age}=0.0027$ is less than $\alpha=0.05$. We reject the null hypothesis. This means there is statistically significant evidence that Age is a good predictor for LNPAID.

```{r}
AutoClaims %>%
  ggplot(aes(x=AGE, y=LNPAID)) +
  geom_point() +
  theme_minimal() +
  labs(title= "Scatter-plot between AGE and LNPAID",
       x= "AGE in years",
       y= "LNPAID in log($$)") +
  geom_smooth(method="lm", se=FALSE, col="blue")
```

While AGE is a statistically significant predictor for LNPAID. The goodness of fit between AGE and LNPAID seems very poor based on $\hat\beta_{Age}$ and $R^2$ values. With $R^2=0.001329$ and $\hat\beta_{Age}=-0.003658$. We can say that there is not a strong relationship between AGE and LNPAID.

#### b.)
```{r}
AutoClaimsRegb <- lm(LNPAID ~ CLASS, data=AutoClaims)

summary(AutoClaimsRegb)

AutoClaimsRegb2 <- lm(LNPAID ~ 0+CLASS, data=AutoClaims)
summary(AutoClaimsRegb2)
```

##### i.)
The point estimate for the LNPAID of claims for the class C7 appears to be $6.940938-0.039775=6.901163$

##### ii.)
```{r}
-0.039775 + c(-1, 1)*-0.748*0.053191
-0.04 + c(-1,1)*-0.75*0.05
```

```{r}
confint.lm(AutoClaimsRegb, 0.95)
```

##### iii.)
```{r}
exp(-0.039775 + c(-1, 1)*-0.748*0.053191)
```

#### c.)
```{r}
AutoClaimsRegc <- lm(LNPAID ~ AGE+GENDER+STATE+CLASS, data = AutoClaims)

summary(AutoClaimsRegc)
anova(AutoClaimsRegc)

AutoClaimsRegc2 <- lm(LNPAID~AGE+GENDER+STATE, data=AutoClaims)
anova(AutoClaimsRegc2, AutoClaimsRegc)
```

##### i.)
$$
H_0: \beta_{Gender}=0 \\
H_A: \beta_{Gender}\not=0
$$
If the F-test of Gender is less than $\alpha=0.05$, then we reject the null hypothesis

Since the F-test yielded a p-value of 0.1759 and is greater than $\alpha=0.05$. We fail to reject the null hypothesis. This means there is no statistically significant evidence that Gender is a good predictor.

##### ii.)
$$
H_0: \beta_{CLASS}=0 \\
H_A: \beta_{CLASS}\not=0
$$

If the F-test of Class is less than $\alpha=0.05$, then we reject the null hypothesis

Since the F-test yielded a p-value of 0.0236 and is less than $\alpha=0.05$. We reject the null hypothesis. This means there is statistically significant evidence that Class is a good predictor.

##### iii.)
```{r}
6.974923-0.003021*60+0.038953+0.098538-0.021854
exp(6.974923-0.003021*60+0.038953+0.098538-0.021854)
```

##### iv.)
The coefficient for someone designated in the class C7 is -0.021854. This means that holding all other variables constant, someone designated the class C7 pays -0.021854 log dollars less than someone designated the class C1.