---
title: "TakeHomeQuiz4.7"
output: pdf_document
editor_options:
  markdown:
    wrap: 72
---

```{r}
library(tidyverse)
library(tinytex)
```

### 4.7
```{r}
UNLifeExpectancy <- read.csv(file.path(str_remove(getwd(), "/Class Notes"), "/Book Data/UNLifeExpectancy.csv"))
```

```{r}
UNLifeExpectancy <- UNLifeExpectancy %>%
  mutate(LNHEALTH = log(PRIVATEHEALTH))
```

#### a.)

$$H_O: \beta_{PUBLICEDUCATION}+\beta_{LNHEALTH}=0$$
$$H_A: \beta_{PUBLICEDUCATION}+\beta_{LNHEALTH}\not=0$$

If the partial F-test p-value is less than $\alpha=0.05$, then we reject the null hypothesis.

```{r}
UNLEregaF <- lm(LIFEEXP~FERTILITY+PUBLICEDUCATION+LNHEALTH, data=UNLifeExpectancy)
UNLEregaR <- lm(LIFEEXP~FERTILITY, data=UNLifeExpectancy)
SSEaF <- sum(residuals(UNLEregaF)^2)
SSEaR <- sum(residuals(UNLEregaR)^2)
s2aF <- SSEaF/df.residual(UNLEregaF)
FRatioA <- ((SSEaR-SSEaF)/(df.residual(UNLEregaR)-df.residual(UNLEregaF)))/(SSEaF/df.residual(UNLEregaF))
FRatioA
1-pf(FRatioA, df.residual(UNLEregaR)-df.residual(UNLEregaF), df.residual(UNLEregaF))
```

Since the p-value from the partial F-test of 0.55 is greater than $\alpha=0.05$. We fail to reject the null hypothesis. This means that there is no statistically significant evidence that PUBLICEDUCATION and LNHEALTH are jointly significant in determining the life expectancy of a country.

#### b.)

```{r}
UNLifeExpectancy %>%
  ggplot(aes(x=REGION, y=LIFEEXP, group=REGION)) +
  geom_boxplot() + 
  theme_gray() +
  labs(title="Boxplot of Region to Life Expectancy",
       x="Region",
       y="Life Expectancy")
```

From the box-plot above, we can see that different regions have their own distribution of life expectancy in their respective countries. However, Region 6 and 8 have medians very different from the other regions. Where their medians are not within the inter-quartile range of the other regions. In other words, it would be safe to say that the region specific means are not similar to each other.

#### c.)

$$H_O: \beta_{REGIONi}=0 \text{ for all i in REGION}$$
$$H_A: \text{At least one }\beta_{REGIONi}\not=0$$
If the F-test p-value is less than $\alpha=0.05$, then we reject the null hypothesis.

```{r}
UNLEregc <- lm(LIFEEXP~as.factor(REGION), data=UNLifeExpectancy)
UNLEregcN <- lm(LIFEEXP~0, data=UNLifeExpectancy)
anova(UNLEregcN, UNLEregc)
```

Since the F-test p-value of essentially 0 is less than $\alpha=0.05$. We reject the null hypothesis. This means there is statistically significant evidence that at least one region is significant in computing a country's life expectancy.

#### d.)
```{r}
UNLEregd <- lm(LIFEEXP~FERTILITY+PUBLICEDUCATION+LNHEALTH+as.factor(REGION), data=UNLifeExpectancy)
```

##### i.)
```{r}
summary(UNLEregd)
CMatdi1 <- matrix(c(1, 2, 5, 1, rep(0,times=7)), nrow=1)
CMatdi2 <- matrix(c(1, 2, 5, 1, rep(0, times = 4), 1, 0, 0), nrow=1)
betaHatd <- coef(UNLEregd)
CMatdi1 %*% betaHatd
CMatdi2 %*% betaHatd
```
The estimated life expectancy for an Arab state with a fertility rate of 2%, public education spending at 5%, and a ln(Health Expenditure) of 1 is ~75 years.

The estimated life expectancy for a sub-Saharan African country with a fertility rate of 2%, public education spending at 5%, and a ln(Health Expenditure) of 1 is ~61 years.

##### ii.)

```{r}
UNLEregdii <- lm(LIFEEXP~0+PUBLICEDUCATION+LNHEALTH+as.factor(REGION), data=UNLifeExpectancy)
summary(UNLEregdii)
CMatdii <- matrix(c(0, 0, 1, rep(0, times=4), -1, 0, 0), nrow=10)
Xdii <- model.matrix(UNLEregdii)
s2d <- sum(residuals(UNLEregdii)^2)/df.residual(UNLEregdii)
se <- sqrt(s2d*t(CMatdii)%*%solve(t(Xdii)%*%Xdii)%*%CMatdii)

t(CMatdii)%*%coef(UNLEregdii)+c(-1,1)*qt(0.975, df.residual(UNLEregdii))*se
```

We are 95% confident that the difference in life expectancy between an Arab State and a Sub-Saharan African Country can be found between 15.4 years and 22.9 years.

##### iii.)
```{r}
CMatdiii <- c(rep(0, times=8), 1, 0 ,-1)
abs(CMatdiii %*% coef(UNLEregd))
```

The point estimate in the difference between a Sub-Saharan African Country and a high-income OECD Country is ~18.2 years.